Comparative Evaluation of Spark and Flink Stream Processing

### 2017-03-05 paper/paper.tex
updated header page

 \thispagestyle{empty}
\centering\LARGE {\bf Comparative Evaluation of Spark and Flink Stream Processing}

\vspace{14pt}

\centering
 Lab: Data Science and Big Data MA-INF 4306,
 Winter Term 2016/2017, University of Bonn

\vspace{14pt}
\centering
 Ehab Qadah\\
 Supervisor: PD Dr. Michael Mock
\rule{\textwidth}{1pt}
\end{flushleft}
\newpage
\tableofcontents

\newpage

@@

\end{abstract}
Recent years have witnessed a advances in the  development of Big Data frameworks like Apache Spark and Apache Flink. For example, both frameworks support real-time
streaming processing. While which platform is the best still an open question. 
In this paper, we provide a performance comparison of streaming data processing in Apache Spark and Apache Flink, by measuring different performance metrics, namely latency and throughput. In addition, we cover some key aspects of streaming data applications and how they are handled in both frameworks. The experiment is done using multiple streaming workloads over real-world datasets.
\section{Introduction}

@@

\par The purpose of this paper is to provide a comparative experimental evaluation of throughput, latency (i.e., performance) of stream processing in Apache Spark and Apache Flink. We use a datasets of airplanes trajectories provided in the context of Datacron project\footnote{\url{http://www.datacron-project.eu/}}, in order to simulate real-world use cases. Using multiple streaming data workloads. Also we try to cover general aspects of streaming processing tasks (later details in Section 3) and how they can be solved in each platform. Informally,  this work aims to help the developers to determine which platform to choose in production for different use cases.

\par The remainder of this paper is organized as follows.
In Section 2, we present the main characteristics, APIs, and main data abstraction of Apache Spark and Flink. Section 3  presents the general aspects of streaming processing, the design and implementation of experimental workloads. Section 4
provides and compares the performance results of the different workloads. Last, Section 5 gives the overall conclusion.

@@

  \centering
    \includegraphics[width=\textwidth, height=.4\textheight]{flink_engine.png}
     \caption{ An example of data flow graph \cite{flink}.}
\end{figure} 
\subsection{Apache Kafka}
\par Apache Kafka is scalable, fault-tolerant distributed publish/subscribe streaming framework \cite{kafka}.
It manges the stream records in different categorizes (i.e., topics) that portioned and distributed over the servers of the Kafka cluster. It allows the data producers to publish stream of records to certain Kafka topic. While the consumer applications can subscribe to one or more topic to read the data stream. The consumers are grouped into groups based on the group name that allows Kafka to distribute and balance the stream partitions among the  members of a certain group for the sake of scalability. Figure 3 shows how 4 partitions of stream is distributed differentially between two groups of consumers based on the number of members. 
Apache Kafka has been widely  adopted, for example, 
Spark and Flink can receive data stream from Kafka.

>>>>

\centering { \large \bf Comparative Evaluation of Spark and Flink Stream Processing}
\vspace{4pt}

\centering
 Lab Report: Data Science and Big Data MA-INF 4306,
 Winter Term 2016/2017, University of Bonn

\vspace{4pt}
\centering
 Ehab Qadah\\
 Supervisor: PD Dr. Michael Mock.
 \vspace{4pt}

 \today
\end{flushleft}

@@

\end{abstract}
Recent years have witnessed a advances in the  development of Big Data frameworks like Apache Spark and Apache Flink. For example, both frameworks support real-time
streaming processing. While which platform is the best still an open question. 
In this report, we provide a performance comparison of streaming data processing in Apache Spark and Apache Flink, by measuring different performance metrics, namely latency and throughput. In addition, we cover some key aspects of streaming data applications and how they are handled in both frameworks. The experiment is done using multiple streaming workloads over real-world datasets.
\section{Introduction}

@@

\par The purpose of this report is to provide a comparative experimental evaluation of throughput, latency (i.e., performance) of stream processing in Apache Spark and Apache Flink. We use a datasets of airplanes trajectories provided in the context of Datacron project\footnote{\url{http://www.datacron-project.eu/}}, in order to simulate real-world use cases. Using multiple streaming data workloads. Also we try to cover general aspects of streaming processing tasks (later details in Section 3) and how they can be solved in each platform. Informally,  this work aims to help the developers to determine which platform to choose in production for different use cases.

\par The remainder of this report is organized as follows.
In Section 2, we present the main characteristics, APIs, and main data abstraction of Apache Spark and Flink. Section 3  presents the general aspects of streaming processing, the design and implementation of experimental workloads. Section 4
provides and compares the performance results of the different workloads. Last, Section 5 gives the overall conclusion.

@@

  \centering
    \includegraphics[width=\textwidth, height=.3\textheight]{flink_engine.png}
     \caption{ An example of data flow graph in Flink \cite{flink}.}
\end{figure} 
\subsection{Apache Kafka}
\par Apache Kafka is scalable, fault-tolerant distributed publish/subscribe streaming framework \cite{kafka}.
It manges the stream records in different categorizes (i.e., topics) that portioned and distributed over the servers of the Kafka cluster. It allows the data producers to publish stream of records to certain Kafka topic. While the consumer applications can subscribe to one or more topic to read the data stream. The consumers are grouped into groups based on the group name that allows Kafka to distribute and balance the stream partitions among the  members of a certain group for the sake of scalability. Figure 3 shows how 4 partitions of stream is distributed differentially between two groups of consumers based on the number of members. 
Apache Kafka has been widely  adopted, for example, 
Spark and Flink can receive data stream from Kafka.
